### 2013-06-19 text/Chapter01.tex
Merge changes

Some intro

\section{How lambda works}
The key difference - how lambda works

Effect systems, introduced by Gifford and Lucassen
\cite{effects-gifford}, track \emph{effects} of computations, such as
memory access or message-based
communication~\cite{effects-messagepassing}. Their approach augments
typing judgments with effect information: $\Gamma \vdash e :
\tau, F$.  Wadler and Thiemann explain how this shapes
effect analysis of lambda abstraction~\cite{monads-effects-marriage}:
%
\begin{quote}
\emph{In the rule for abstraction, the effect is empty because evaluation immediately
returns the function, with no side effects. The effect on the function arrow
is the same as the effect for the function body, because applying the function will
have the same side effects as evaluating the body.}
\end{quote}
%
In contrast to the static analysis of \emph{effects}, the analysis of \emph{context-dependence} 
does not match this pattern. In the systems we consider, lambda abstraction places 
requirements on both the \emph{call-site} (latent requirements) and the \emph{declaration-site} 
(immediate requirements), resulting in different syntactic properties. 
We informally discuss three examples first that demonstrate how contextual
requirements propagate. Section ? then unifies these 
in a single calculus.

Also \cite{effects-producer-semantics}

\begin{quote}
\emph{We will define an effect as a producer effect if all computations with that effect 
can be thunked as "pure" computations for a domain-specific notion of purity.}
\end{quote}

Demonstrate using distributed computations

Demonstrate using dataflow/liveness

Perhaps mention LINQ as a motivation for cross-compilation..
\cite{app-linq}

\section{Associating with context}

-> Type providers make this important

@@

\section{Flat coeffect system}
~

resources?

@@

\section{Structural coeffect system}
\label{sec:structural}

The \emph{flat coeffect system} presented in the previous sections has a number of uses, but often
we need to track context-dependence in a more fine-grained way. To track neededness or security, 
we need to associate information with individual \emph{variables} of the context.

At the same time, we want to avoid developing multiple variants of coeffect systems -- indeed, 
our motivation is to develop a \emph{single unified mechanism} for tracking context-dependence.
In this section, we present a more powerful \emph{structural coeffect calculus}, which is a 
generalization of the flat calculus.

%We show that \clflt~can be obtained as a special case of \clstr. The
%structural version of our system is also where our work differs more significantly from well-known
%effect systems.

% --------------------------------------------------------------------------------------------------

\subsection{Motivation: Tracking array accesses}
\label{sec:structural-motivation}
\newcommand{\pastval}[2]{\ident{#1}_{[#2]}}
\newcommand{\dnat}{\ident{D}_\ident{nat}}
\newcommand{\cprd}{\times}
\newcommand{\cvop}{~\ident{max}~}

Similarly to the flat version, the \emph{structural coeffect calculus} works with contexts and
functions annotated with a coeffectt tags, written $\ctyp{r}{\Gamma}$ and $\ctyp{r}{\tau_1} \rightarrow \tau_2$, 
respectively, but we use richer tag structure.

As an example, consider a language that allows us to get a value of a variable (representing
some changing data-source) \ident{x} versions back using the syntax $\pastval{a}{x}$. 
To track information about individual variables, we use a product-like operation $\cprd$ on tags 
to mirrors the product structure of variables. For example:
%
\begin{equation*}
\begin{array}{l}
\ctyp{5 \cprd 10}{(\ident{a}:\dnat, \ident{b}:\dnat)}
  \vdash
    \pastval{a}{5}+\pastval{b}{10}: \ident{nat}
\end{array}
\end{equation*}
%
The coeffect tag $5 \cprd 10$ corresponds to the free-variable context $\ident{a}, \ident{b}$, denoting
that we need at most 5 and 10 past values of \ident{a} and \ident{b}. If we substitute \ident{c}
for both \ident{a} and \ident{b}, we need another operation to combine multiple tags associated with 
a single variable:
%
\begin{equation*}
\begin{array}{l}
\ctyp{5 \cvop 10}{(\ident{c}:\dnat)}
  \vdash
    \pastval{c}{5}+\pastval{c}{10}: \ident{nat}
\end{array}
\end{equation*}
%
In this example, the operation $\cvop$ would be the \emph{max} function and so $5 \cvop 10 = 10$.
Before looking at the formal definition, consider the typing of let bindings:
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{c} = \kvd{if}~\ident{test}()~\kvd{then}~\ident{a}~\kvd{else}~\ident{b}\\
\pastval{a}{15}+\pastval{c}{10}
\end{array}
\end{equation*}
%
The expression has free variables \ident{a} and \ident{b} (we ignore \ident{test}, which
is not a data source). It defines \ident{c}, which may be assigned either \ident{a} or \ident{b}.
The variable \ident{a} may be used directly (second line) or indirectly via \ident{c}.

The expression assigned to \ident{c} uses variables \ident{a} and \ident{b}, so its typing
context is $\ctyp{0 \cprd 0}{(\ident{a}, \ident{b})}$. The value $0$ is the unit
of $\cvop$ and it denotes empty coeffect. The typing context of the
body is $\ctyp{15 \cprd 10}{(\ident{a}, \ident{c})}$.

To combine the tags, we take the coeffect associated with \ident{c} and apply it to the tags
of the context in which \ident{c} was defined using the $\cvop$ operation. This is then combined 
with the remaining tags from the body yielding the overall context:
$\ctyp{15 \cprd (10 \cvop (0 \cprd 0))}{(\ident{a}, (\ident{a}, \ident{b}))}$. Using a simple
normalization mechanism (described later), this can be further reduced to
$\ctyp{(15 \cvop 10) \cprd 10}{(\ident{a}, \ident{b})}$. This gives us the required information --
we need at most $\emph{max}(15,10)$ past values of \ident{a} and at most $10$ past values of 
\ident{b}.

@@

>>>>

Modern software applications behave differently depending on the environment or context 
in which they execute. They often run in increasingly rich environments that provide 
resources (e.g. database or GPS sensor) and are gradually more diverse (e.g. multiple 
versions of different mobile platforms). Web applications are split between client, server 
and mobile components; mobile applications must be aware of the context and of the platform 
while the ``internet of things'' makes the environments even more heterogeneous; 
applications that access rich data sources need to propagate security policies and provenance 
information about the data.

Writing such context-aware applications is a fundamental problem of modern software 
engineering. The state of the art relies on ad-hoc approaches -- using hand-written conditions 
or pre-processors for conditional compilation. Common problems that developers face include:

\begin{itemize}
\item \textbf{System capabilities.} When writing code that is cross-compiled to multiple 
  targets (e.g. SQL, CUDA or JavaScript) the compilation often occurs at runtime and developers 
  have no guarantee that it will succeed until the program is executed.

\item \textbf{Platform versioning.} When developing application for multiple versions of a 
  system (e.g. Android), developers rely on lazy loading at runtime or use conditional compilation 
  using \ident{\#if}. The former delays errors to runtime, while the latter requires building all 
  possible configurations to discover simple compile-time errors.

\item \textbf{Resources \& data availability.} When programming applications that access 
  resources or data provided by the environment (e.g. specific database table, GPS sensor), 
  the program typically performs dynamic check for the resource availability. However, this 
  is not checked in any way -- we have no easy way to tell what happens when the resource is 
  not available (e.g. is there a fallback strategy or not?)

\item \textbf{Meta-programming.} In distributed client-server programming, we often need to 
  manipulate code using meta-programming techniques on the server and then pass a closed 
  expression to client, translated to another language (JavaScript). We need to be able to 
  guarantee that such expressions are well-formed before they are translated.
\end{itemize}

The issues discussed above pose an important problem to current and future state of computing.
In this thesis, we develop a unifying mathematical model that allows us to capture the properties
discussed above. We use the model in three different ways to capture different notions of 
context-dependence.

In the rest of the chapter, we first look at the distinguishing factor of two of the presented
coeffect systems (how they differ from other existing work). Then we look at a number of concrete 
problems that are discussed in greater details in later chapters. We use the examples to demonstrate 
the real-world problems that motivate our work. (Theoretical and other motivations are discussed 
later in Chapter~\ref{ch:related-work}).

\section{Context and lambda abstraction}

Our work on context-aware programming languages connects two directions in existing research on the 
theory of programming languages. On one side, effect systems \cite{effects-gifford} and monadic 
computations \cite{monad-notions,monads-effects-marriage} provide a detailed and established method 
for tracking what effects programs have -- that is, how they affect the environment where they execute. 
On the other side, the work on comonadic notions of computations \cite{comonads-notions} shows how 
to use the mathematical dual of monads -- comonads -- to give categorical semantics of 
context-dependent computations.

Effect systems introduced track actions such as memory operations or communication. They are described 
by typing judgments of a form $\Gamma \vdash e : \alpha, \sigma$ where $\Gamma$ is the context of a 
program (typically available variables), $e$ is the expression (program) itself, $\alpha$ is the type of 
values returned by the program (e.g. integer or boolean) and $\sigma$ is a set of possible effects. The 
judgment states that, given the context $\Gamma$, an expression has a type $\alpha$ and can only perform 
effects specified by the set $\sigma$. Wadler and Thiemann \cite{monads-effects-marriage} explain how 
this shapes effect analysis of a lambda abstraction -- that is, how effect systems analyze the effects 
associated with a definition of a function:
%
\begin{quote} 
\emph{In the rule for abstraction, the effect is empty because evaluation immediately
returns the function, with no side effects. The effect on the function arrow
is the same as the effect for the function body, because applying the function will
have the same side effects as evaluating the body.}
\end{quote}
%
This means that, when a programmer defines a function, the system records that executing the 
function will perform the effects of the body of the function. However, simply defining a function
is an effect-free computation. Tate \cite{effects-producer-semantics} calls such effect systems
\emph{producer} effect systems and generalizes the idea of function to more general ``thunking'':
%
\begin{quote}
\emph{We will define an effect as a producer effect if all computations with that effect 
can be thunked as "pure" computations for a domain-specific notion of purity.}
\end{quote}
%
In contrast to the static analysis of (producer) effect systems, the analysis of 
\emph{context-dependence} does not match this pattern. In the systems we consider, lambda 
abstraction places requirements on both the \emph{call-site} (latent requirements) and the 
\emph{declaration-site} (immediate requirements), resulting in different syntactic properties. 
We informally discuss three examples first that demonstrate how contextual requirements propagate.

\section{Why coeffects matter}

This section gives three examples of context-dependent computations whose properties can be captured
by the tree calculi presented in this thesis. We look at an example of the \emph{flat coeffect calculus} 
(Chapter~\ref{ch:flat-coeffects}), \emph{structural coeffect calculus} (Chapter~\ref{ch:structural-coeffects})
and \emph{coeffect meta-language} (Chapter~\ref{ch:coeffect-metalanguage}).

@@

\subsection{Flat coeffect calculus}

The flat coeffect calculus associates a single piece of information with the context. As an example,
we look at a simple distributed programming language that includes the concept of \emph{resources}.
A resource may be accessed using a special construct $\kvd{access}~\ident{Res}$. The following 
example shows a function taht lists recent events -- it accesses the resource \ident{News} 
(representing a database) and a resource \ident{Clock} (with the current time):
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{recentEvents} = \lambda () \rightarrow\\
\quad\kvd{let}~\ident{db} = \kvd{access}~\ident{News}~\kvd{in}\\
\quad\ident{query}~\ident{db}~\str{"SELECT * WHERE Date > \%1"}~(\kvd{access}~\ident{Clock})
\end{array}
\end{equation*}
%
Consider a scenario where the function is \emph{declared} on the server-side and then
transferred and \emph{executed} on the client-side. The resource \ident{News} represents a 
database that is only available on the server-side and so the function needs to keep a remote
reference to the server. However, the \ident{Clock} resource may (or may not) be available on the
client-side. If the resource is available on the client, then it may be re-bound and the function
will use the current client's data -- for example, to accommodate for time-zone changes.

This example demonstrates how lambda abstraction behaves for context-dependent computations. The 
context requirement of the function body is a set of resources $\{\ident{Clock}, \ident{News}\}$.
The context requirements are split between the \emph{declaration-site} and \emph{call-site}.
However, there are multiple possible splittings. The splitting $\{\ident{News}\} \cup \{\ident{Clock}\}$
models the case when the database is accessed from the server, but time is taken from the client,
while the splitting $\{\ident{Clock}, \ident{News}\} \cup \{\}$ corresponds to the case when both
resources are accessed from the server.

@@

\subsection{Structural coeffect calculus}

\newcommand{\pastval}[2]{\ident{#1}_{[#2]}}
\newcommand{\dnat}{\ident{stream}}

The calculus used in the previous section annotates the entire context with a single value --  such
as the set of required resources. However, sometimes it is desirable to annotate not the entire 
context, but individual variables of the context.

As an example, consider a language that allows us to get a value of a variable (representing
some changing data-source) \ident{x} versions back using the syntax $\pastval{a}{x}$. 
To track information about individual variables, we use a product-like operation $\times$ on tags 
to mirrors the product structure of variables. For example:
%
\begin{equation*}
\begin{array}{l}
(\ident{a}:\dnat, \ident{b}:\dnat) ~@~ 5 \times 10
  \vdash
    \pastval{a}{5}+\pastval{b}{10}: \ident{nat}
\end{array}
\end{equation*}
%
The annotations $5 \times 10$ corresponds to the free-variable context $\ident{a}, \ident{b}$, denoting
that we need at most 5 and 10 past values of \ident{a} and \ident{b}, respectively. If we substitute 
\ident{c} for both \ident{a} and \ident{b}, we need to combine the context-requirements and take the
maximum of the requirements of the individual variables:
%
\begin{equation*}
\begin{array}{l}
(\ident{c}:\dnat) ~@~ \ident{max}(5, 10)
  \vdash
    \pastval{c}{5}+\pastval{c}{10}: \ident{nat}
\end{array}
\end{equation*}
%
This version of the calculus removes the non-determinism of the lambda abstraction from the previous
version. As we associate information with individual variables, lambda abstraction creates a function
that requires the context requirements associated with the variable that is being abstracted over.

For example, if we wrapped the earlier example above in a function taking \ident{a} (and using 
\ident{b} from the declaration-site) then the function would have context requirements $5$ -- that
is the number associated with the variable \ident{a}.

\subsection{Coeffect meta-language}

\todo{Give some example that uses the coeffects metalanguage style. This will probably be 
meta-programming with open expressions (similarly to what Pfenning and Nanevski do with 
contextual modal type theory).}

\section{Why context matters}

We claimed earlier that context-dependent comptuations are becoming increasingly common and our
work is focused on annotating the (free-variable) context with additional information.
The importance of context can be also demonstrated by looking at a technology that focuses 
solely on making the context richer -- F\# type providers.

\todo{Say more about type providers -- they extend the context $\Gamma$ so that it can be 
lazily loaded and it can be huge. Potentially, it could be also annotated with additional
meta-data...}

@@

\todo{We present three calculi that model common notions of context-dependence and can be used
as basis for developing context-aware programming langauges with static type systems.}

### 2013-06-19 text/Chapter03.tex
Refactor rules
Merge changes

\chapter{Flat coeffects} 
\label{ch:flat-coeffects}

@@

TODO: Change this

@@

>>>>

\chapter{Flat coeffect language} 
\label{ch:flat-coeffects}

@@

\todo{Update the introduction \& edit the whole chapter -- currently, this is just a direct copy
from the accepted ICALP paper on flat coeffects.}

\todo{Other extensions -- add more examples of indexed comonoads, consider recursion and 
fixed-point operator, possibly also consider partially defined operations, which lets us
express type systems that rule out some computations (rather than just static analyses). }

@@

\newcommand{\cprd}{\times}
\newcommand{\cvop}{\oplus}

### 2013-06-19 text/Chapter04.tex
Merge changes

\chapter{Structural coeffect language} 
\label{ch:structural}

@@

Introduction

@@

\section{Introduction}
\label{sec:structural-intro}

We already talked about tainting in pathways to coeffects... here is more stuff!

~

~

Our second motivating example is the use of \emph{tainting} for the prevention of SQL injection 
attacks \cite{app-secure-tainting}. The idea is to (statically) check that SQL commands cannot be 
directly constructed from, potentially dangerous, inputs provided by the user. Consider the type
checking of the following expression in a context containing variables \ident{id} and \ident{msg}:
%
\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{name} = \ident{query}~(\str{"SELECT Name WHERE Id = "}~+~\ident{id})~\kvd{in}\\
\ident{msg}~+~\ident{name}
\end{array}
\end{equation*}
%
We say that a variable is \emph{tainted} if it may come from untrusted input. In the above example,
\ident{id} must not be tainted. Otherwise, the attacker could specify values such as 
\str{"1; DROP TABLE Users"}. The variable \ident{msg} may or may not be tainted, because it is not
used in protected context (i.e.~to construct an SQL query). Using the coeffect calculus, security
information can be naturally attached to the context. We use tag \ident T for tainted and 
\ident N for not tainted:
%
\begin{equation*}
\ctyp{\ident T \times \ident N}{(\ident{id}:\ident{string}, \ident{msg}:\ident{string})} 
  \vdash (\ldots) : \ident{string}
\end{equation*}
%
The tags \ident T and \ident N correspond to variables \ident{id} and \ident{msg}, respectively.
The checking of tainting is a special case of checking of \emph{non-interference} in 
\emph{secure information flow} \cite{app-secure-information-flow}. Languages such as FlowCaml 
\cite{app-security-flowcaml} can check such properties statically, but it is also possible to 
embed such checking in a monadic library \cite{monad-secure-flow}.

Assuming the monadic type $\mtyp{s}{\tau}$ represents a value of type $\tau$ 
in a security context $s$ (either \ident T or \ident N): 
%
\begin{equation*}
\ident{id} : \mtyp{\ident T}{\ident{string}}, \ident{msg} : \mtyp{\ident N}{\ident{string}} 
  \vdash (\ldots) : \mtyp{\ident T}{\ident{string}}
\end{equation*}
%
In this example, the result is tagged with \ident T to denote that it may be tainted.
If the result was tagged with \ident N, the first parameter would also have to be \ident N
(and we would get fewer information about the function).

The monadic encoding has an unfortunate property from the semantic perspective. 
The function has monadic structure over the \emph{domain} as well as over the \emph{codomain}
and thus it does not match the desirable Kleisli model of expressions $\tau_1 \times \ldots \tau_n \rightarrow \mtyp{}{\tau}$. 
This makes the it difficult to give standard composable semantics of the library.

In coeffect calculus, expressions are modelled using the coKleisli form 
$\ctyp{r_1 \times \ldots \times r_n}{(\tau_1 \times \ldots \tau_n)} \rightarrow \tau$ with a tags 
$r_n$ that specify taintedness of inputs. As discussed in detail in Section~\ref{sec:structural-tainting}, 
this model gives a simple comonadic semantics of language-based security using  tainting.

%---------------------------------------------------------------------------------------------------
%===================================================================================================

\begin{figure*}[t]
% Constants, if we wanted them, would look like this
%
% \inference[]
%   {c : \tau \in \Phi}
%   {\ctyp{1}{()} \vdash c : \tau}&
% \textnormal{\rulest{const}}
% \\[1.6em]

%\paragraph{Typing rules}
\begin{equation*}
\boxed{\ctyp{r}{\Gamma} \vdash e : \tau }
\end{equation*}

\begin{equation*}
\begin{array}{cl}

\inference[]
  {}
  {\ctyp{0}{(v:\tau)} \vdash v : \tau }&
\textnormal{\rulest{var}}
\\[1.8em]
\inference
  {\ctyp{r \csprd s}{(\Gamma, v:\tau_1)} \vdash e : \tau_2 & v \notin \Gamma}
  {\ctyp{r}{\Gamma} \vdash \lambda v.e : \ctyp{s}{\tau_1} \rightarrow \tau_2 }&
\textnormal{\rulest{fun}}
\\[1.8em]
\inference
  {\ctyp{s}{\Gamma_1} \vdash e_1 : \ctyp{r}{\tau_1} \rightarrow \tau_2 &
   \ctyp{t}{\Gamma_2} \vdash e_2 : \tau_1 }
  {\ctyp{s \csprd (r \csvop t)}{(\Gamma_1, \Gamma_2)} \vdash e_1~e_2 : \tau_2}&
\textnormal{\rulest{app}}
\\[1.8em]
\inference
  {\ctyp{r \csprd s}{(\Gamma_1, v:\tau_1)} \vdash e_1 : \tau_2 &
  \ctyp{t}{\Gamma_2} \vdash e_2 : \tau_1}
  {\ctyp{r \csprd (s \csvop t)}{(\Gamma_1, \Gamma_2)} \vdash \kvd{let}~x=e_2~\kvd{in}~e_1 : \tau_2}&
\textnormal{\rulest{let}}
\\[1.8em]
\inference
  {\ctyp{r}{\Gamma} \vdash e : \tau &
   \ctxtrans{\ctyp{r'}{\Gamma'}}{\ctyp{r}{\Gamma}}}
  {\ctyp{r'}{\Gamma'} \vdash e : \tau}&
\textnormal{\rulest{ctx}}
\end{array}
\end{equation*}

%
% Empty variable context corresponds to unit, but things get messy
%

%\paragraph{Context rules}
\begin{equation*}
\boxed{\ctxtrans{\ctyp{r}{\Gamma}}{\ctyp{r}{\Gamma}} }
\end{equation*}

\begin{equation*}
\begin{array}{rcll}
\multicolumn{3}{c}{
  \inference
    {  \ctxtrans{\ctyp{r'}{\Gamma_1'}}{\ctyp{r}{\Gamma_1}} }
    {  \ctxtrans{\ctyp{r'\csprd s}{(\Gamma_1', \Gamma_2)}}{\ctyp{r\csprd s}{(\Gamma_1, \Gamma_2)}} }
}&
\textnormal{\rulest{nest}}
\\[2.0em]
  \ctxtrans{\ctyp{r\csprd s}{(\Gamma_1, \Gamma_2)} &}{& \ctyp{s\csprd r}{(\Gamma_2, \Gamma_1)}}&
\textnormal{\rulest{exch}}
\\[0.8em]
  \ctxtrans{ \ctyp{r \csprd 1}{(\Gamma, ())} &}{&
    \ctyp{r}{\Gamma}}&
\textnormal{\rulest{empty}}
\\[0.8em]
  \ctxtrans{\ctyp{r\csprd 0}{(\Gamma, x:\tau_1)} &}{& \ctyp{r}{\Gamma}}&
\textnormal{\rulest{weak}}
\\[0.8em]
  \ctxtrans{\ctyp{r \csvop s}{(x:\tau)} &}{&
    \ctyp{r \csprd s}{(x:\tau, x:\tau)}}&
\textnormal{\rulest{contr}}
\\[0.8em]
  \ctxtrans{\ctyp{r\csprd (s\csprd t)}{(\Gamma_1, (\Gamma_2, \Gamma_3))} &}{&
    \ctyp{(r\csprd s)\csprd t}{((\Gamma_1, \Gamma_2), \Gamma_3)}}\;&
\textnormal{\rulest{assoc}}
\\[0.8em]
\multicolumn{3}{c}{
  \ctxtrans{\ctyp{r}{\Gamma}}{\ctyp{s}{\Gamma}}\quad(\textnormal{when}~s \csvop r = s)
}&
\textnormal{\rulest{sub}}
\end{array}
\end{equation*}

% We do not need this rule (can be expressed in terms of the other one)
%
% \inference
%  {  \ctxtrans{\ctyp{s}{\Gamma_2}}{\ctyp{s'}{\Gamma_2'}} }
%  {  \ctxtrans{\ctyp{r\csprd s}{(\Gamma_1, \Gamma_2)}}{\ctyp{r\csprd s'}{(\Gamma_1, \Gamma_2')}} }&
% \textnormal{\rulest{right}}

\caption{Type system for \clstr}
\label{fig:structural-coeffect-typing}
\end{figure*}

\newpage

\section{Structural coeffect system}
\label{sec:structural}

At the same time, we want to avoid developing multiple variants of coeffect systems -- indeed, 
our motivation is to develop a \emph{single unified mechanism} for tracking context-dependence.
In this section, we present a more powerful \emph{structural coeffect calculus} \clstr, which is a 
generalization of \clflt. We show that \clflt~can be obtained as a special case of \clstr. The
structural version of our system is also where our work differs more significantly from well-known
effect systems.

@@

@@

\subsection{Example: Neededness analysis}
\label{sec:structural-neededness}

\subsection{Example: Tracking tainting}
\label{sec:structural-tainting}

\section{Summary}
Oops!

>>>>

\chapter{Structural coeffect language} 
\label{ch:structural-coeffects}

@@

@@

\section{Introduction}

@@

\ExecuteMetaData[rules/rules.tex]{scl-types}

@@

\subsection{Example: Liveness analysis}
\label{sec:structural-neededness}

\subsection{Example: Data-flow (revisited) }
\label{sec:structural-tainting}

\todo{Also, consider additional language features that we consider for flat coeffects
  (mainly recursion and possibly conditionals)}

\section{Conclusions}

\todo{ (...) }

### 2013-06-19 text/Chapter05.tex
Merge changes

\chapter{Coeffect meta-language} 
\label{ch:metalanguage}

@@

Introduction

@@

?
\ExecuteMetaData[rules/rules.tex]{cml-types}

\section{Summary}
Oops!

>>>>

\chapter{Coeffect meta-language} 
\label{ch:coeffect-metalanguage}

@@

Both flat coeffect calculus and structural coeffect calculus (presented in the past two chapters)
use indexed comonads to define the semantics of the langauge. In this section, we follow the 
meta-language style and embed indexed comonads into the language -- the type constructor
$\ctyp{r}{\alpha}$ becomes a first-class value and we add language constructs corresponding to
primitive operations of the indexed comonad.

@@

\section{Type system}
\ExecuteMetaData[rules/rules.tex]{cml-types}

\section{Operational properties}

\section{Categorical semantics}

\section{Applications}

\subsection{Meta-programming}

\subsection{Mobile computations}

\section{Related work}
This chapter is closely related to Contextual Modal Type Theory (CMTT) of Nanevski et al. However
they develop their language using model logic as a basis, while we use categorical foundations 
as the basis - leading to a different system.

\section{Summary}
